How to select predictive models for decision-making or causal inference?

### 2023-12-08 how_to_select_causal_models.tex
shorter overlap measurement

%\idea{Overlap is important but not easily measurable}
Good overlap between treated and control population is
crucial for causal inference (positivity assumption
\ref{assumption:overlap}). It is often assessed by
comparing visually population distributions as in Figure
\ref{fig:toy_example} or computing standardized difference on each
feature
\cite{austin_introduction_2011,austin_moving_2015}. While these methods are
useful to decide if positivity holds, they do not yield a
single measure. Rather, we compute the divergence
between the population covariate distributions $\mathbb P(X|A=0)$ and $\mathbb
    P(X|A=1)$
\cite{damour_overlap_2020,johansson2022generalization}. We introduce the
Normalized Total Variation (NTV), a divergence based on the sole propensity
score (Appendix \ref{apd:motivation_ntv}).

@@

\paragraph{Motivation of the Normalized Total Variation}
%\idea{A simplier measure, NTV seems to capture what we want}
Computing overlap when working only on samples of the observed distribution,
outside of simulation, requires a sophisticated estimator of discrepancy
between distributions, as two data points never have the same exact set of
features. Maximum Mean Discrepancy \cite{gretton2012kernel} is typically
used in the context of causal inference
\cite{shalit_estimating_2017,johansson2022generalization}. However it
needs a kernel, typically Gaussian, to extrapolate across neighboring
observations. We prefer avoiding the need to specify such a kernel, as it must
be adapted to the data which is tricky with categorical or non-Gaussian
features, a common situation for medical data.

>>>>

%\idea{Overlap is important but not easily measurable}
Good overlap between treated and control population is crucial for causal
inference (positivity assumption \ref{assumption:overlap}). We introduce the
Normalized Total Variation (NTV), a divergence based on the propensity score
summarizing the overlap between both populations (Appendix
\ref{apd:motivation_ntv}).

@@

\paragraph{Motivation of the Normalized Total Variation}

%\idea{A simplier measure, NTV seems to capture what we want}

Overlap is often assessed by comparing visually population distributions as in
Figure \ref{fig:toy_example} or computing standardized difference on each
feature \cite{austin_introduction_2011,austin_moving_2015}. While these methods
are useful to decide if positivity holds, they do not yield a single measure.
Rather, we compute the divergence between the population covariate distributions
$\mathbb P(X|A=0)$ and $\mathbb P(X|A=1)$
\cite{damour_overlap_2020,johansson2022generalization}. Computing overlap when
working only on samples of the observed distribution, outside of simulation,
requires a sophisticated estimator of discrepancy between distributions, as two
data points never have the same exact set of features. Maximum Mean Discrepancy
\cite{gretton2012kernel} is typically used in the context of causal inference
\cite{shalit_estimating_2017,johansson2022generalization}. However it needs a
kernel, typically Gaussian, to extrapolate across neighboring observations. We
prefer avoiding the need to specify such a kernel, as it must be adapted to the
data which is tricky with categorical or non-Gaussian features, a common
situation for medical data.
